\documentclass[a4paper,12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{float}
\geometry{a4paper, margin=2cm}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=left,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2
}
\lstset{style=mystyle}

\title{\textbf{PROJECT REPORT}\\[0.3cm]\Large Topic 2: Creative Project Portfolio Platform\\[0.5cm]\large Focus on MongoDB Database Design and Optimization}
\author{}
\date{\today}

\begin{document}
\maketitle
\tableofcontents
\newpage

%% === SECTION 1: INTRODUCTION ===
\section{Introduction}

\subsection{Background}
In creative project management, the core requirement is storing and retrieving complex hierarchical structures: \textbf{Project {\textrightarrow} Phase (Category) {\textrightarrow} Task {\textrightarrow} Sub-task / Attached Resource}. Traditional Relational Database Management Systems (RDBMS) solve this by creating multiple separate tables (projects, phases, tasks, subtasks) and connecting them via multi-level \texttt{JOIN} operations, resulting in significant query overhead as the data tree grows deeper.

This project builds a system called \textbf{Football Match Planner} --- a platform for managing amateur football match organization, where each ``match'' serves as a ``creative project'' containing a complete 4-level hierarchical structure. The system uses \textbf{MongoDB} as the primary database, leveraging the \textit{Nested Document} model to encapsulate the entire data tree within a single JSON document.

\subsection{Technology Stack}
\begin{itemize}
    \item \textbf{Backend:} FastAPI (Python) --- high-performance API framework with async support.
    \item \textbf{Database (Hybrid Architecture):} 
    \begin{itemize}
        \item \textbf{MongoDB 7.x:} Document database for managing complex, hierarchical nested structures (Matches/Projects).
        \item \textbf{PostgreSQL 15+:} Relational database for structured, ACID-compliant data (Members/Authentication).
    \end{itemize}
    \item \textbf{Drivers:} PyMongo (MongoDB) and SQLAlchemy/psycopg2 (PostgreSQL).
    \item \textbf{Validation:} Pydantic --- application-level input schema definition.
    \item \textbf{Frontend:} Vanilla HTML/CSS/JS with Kanban + Tree Hybrid Editor interface.
\end{itemize}

\subsection{Report Objectives}
This report provides an in-depth analysis of the database aspects:
\begin{enumerate}
    \item \textbf{Hybrid Architecture (Polyglot Persistence):} The combination of PostgreSQL and MongoDB to optimize distinct data patterns.
    \item \textbf{Theoretical Foundation:} Analysis of CAP Theorem, ACID vs. BASE semantics.
    \item \textbf{Schema Design Trade-offs:} In-depth analysis of Embedding vs. Referencing.
    \item Nested Document schema design for a 4-level hierarchical structure.
    \item Advanced Aggregation Pipelines: member performance scoring and automatic Data Cleansing.
    \item \textbf{MongoDB vs SQL Comparison:} Evaluating Data Locality, Read/Write Amplification, and Impedance Mismatch.
\end{enumerate}

\newpage

%% === SECTION 2: THEORETICAL BACKGROUND ===
\section{Theoretical Background: Hybrid Architecture}

\subsection{Polyglot Persistence (MongoDB + PostgreSQL)}
Modern applications encounter diverse data access patterns that a single database paradigm cannot efficiently handle. This project implements a \textbf{Polyglot Persistence} architecture:
\begin{itemize}
    \item \textbf{PostgreSQL (Relational):} Used for managing \texttt{members} (users, authentication, roles). User data is highly structured, requires strong consistency, and rarely involves deep hierarchies.
    \item \textbf{MongoDB (NoSQL):} Used for managing \texttt{matches} (projects, tasks, subtasks). Match data is highly hierarchical (4 levels deep), schema-flexible, and read-heavy in its entirety.
\end{itemize}

\subsection{CAP Theorem and ACID vs BASE}
According to the \textbf{CAP Theorem}, a distributed data store can only guarantee two of three properties: Consistency (C), Availability (A), and Partition Tolerance (P).
\begin{itemize}
    \item \textbf{PostgreSQL (ACID - CA/CP):} Prioritizes Atomicity, Consistency, Isolation, and Durability. It ensures that any transaction involving user credentials or states is strictly consistent immediately.
    \item \textbf{MongoDB (BASE - CP):} Operates on Basically Available, Soft state, Eventual consistency semantics (though modern MongoDB supports ACID transactions). It focuses on high availability and partition tolerance, making it ideal for high-throughput task updates where strict millisecond-level consistency across different nodes is less critical than rapid availability.
\end{itemize}

%% === SECTION 3: METHODOLOGY ===
\section{Methodology and System Implementation}

\subsection{Database Connection}

The file \texttt{db/database.py} and \texttt{db/postgres.py} establish the hybrid connection:

\begin{lstlisting}[language=Python, caption=Hybrid Connection Setup]
# db/database.py (MongoDB for Matches)
from pymongo import MongoClient
MONGO_URI = "mongodb://localhost:27017/"
mongo_client = MongoClient(MONGO_URI)
matches_collection = mongo_client["football_planner"]["matches"]

# db/postgres.py (PostgreSQL for Members)
from sqlalchemy import create_engine
PG_URI = "postgresql://user:pass@localhost:5432/football_planner"
pg_engine = create_engine(PG_URI)
\end{lstlisting}

The system uses \textbf{two distinct databases}:
\begin{itemize}
    \item \texttt{football\_planner.matches} (MongoDB): Stores all project (match) data as nested documents.
    \item \texttt{members} (PostgreSQL): Stores user information with SHA-256 hashed passwords in a relational table.
\end{itemize}

The parameter \texttt{serverSelectionTimeoutMS=5000} ensures the application fails fast (5 seconds) if the MongoDB server is unavailable, instead of hanging indefinitely.

\subsection{Schema Design --- Nested Document Model}

\subsubsection{Embedding vs Referencing Trade-off Analysis}
When designing NoSQL schemas, the primary decision is between Embedding (Nested Documents) and Referencing (Normalized).

\begin{table}[H]
\centering
\begin{tabular}{|l|p{6cm}|p{6cm}|}
\hline
\textbf{Criteria} & \textbf{Embedding (Nested)} & \textbf{Referencing (Normalized)} \\ \hline
\textbf{Relationship} & 1-to-Few (e.g., Match has $<100$ Tasks) & 1-to-Many / Unbounded (e.g., Users to Logins) \\ \hline
\textbf{Read Pattern} & Queried together. Read-heavy. & Queried separately. \\ \hline
\textbf{Data Locality} & \textbf{High.} Fetched in a single disk block. & \textbf{Low.} Requires multiple index scans. \\ \hline
\textbf{Data Mutation} & Infrequent or localized updates. & Frequently updated shared data. \\ \hline
\end{tabular}
\caption{Embedding vs Referencing Trade-offs}
\end{table}

For the \texttt{matches} collection, we chose \textbf{Embedding} because the entire task tree is always loaded simultaneously when viewing the Kanban board, and the number of subtasks per match is strictly bounded.

\subsubsection{Deep Comparison: MongoDB vs SQL for Hierarchical Data}
Using a traditional RDBMS for a 4-level hierarchy involves significant drawbacks:
\begin{itemize}
    \item \textbf{Impedance Mismatch:} In Python, the data is manipulated as nested dictionaries/objects. SQL requires an ORM layer to map flat rows to nested objects, consuming CPU cycles. MongoDB's JSON-like BSON format maps directly to Python dictionaries with zero friction.
    \item \textbf{Read Amplification \& Data Locality:} SQL requires 4 \texttt{JOIN} tables (\texttt{matches}, \texttt{categories}, \texttt{tasks}, \texttt{subtasks}). The database must scan multiple scattered disk pages to assemble the tree. MongoDB retrieves the entire match document from a single contiguous disk sector (High Data Locality), reducing disk I/O.
    \item \textbf{Write Amplification:} SQL excels here; updating a single subtask only touches one row. MongoDB requires rewriting the document (though Array Filters mitigate this by modifying elements in place).
\end{itemize}

\subsubsection{Detailed Schema for the \texttt{matches} Collection}

\begin{lstlisting}[language=json, caption=Complete JSON structure of a document in the matches collection]
{
  "_id": "match_abcd1234",
  "name": "Friendly Match Round 1",
  "date": "2026-05-10",
  "date_end": "2026-05-10T20:00",
  "status": "Planned",
  "created_at": "2026-05-02T10:00:00",
  "categories": [
    {
      "name": "Logistics & Venue",
      "tasks": [
        {
          "name": "Book the pitch",
          "status": "Todo",
          "task_type": "logistics",
          "assigned_to": "Duong",
          "assignee_id": "mem_abc123",
          "location": "Thong Nhat Stadium",
          "cost": 50,
          "subtasks": [
            {"name": "Find suitable pitch", "cost": 0, "status": "Todo"},
            {"name": "Pay deposit", "cost": 10, "status": "Done"}
          ]
        }
      ]
    }
  ],
  "activity_logs": [
    {
      "action": "Created new match: Round 1",
      "timestamp": "2026-05-02T10:00:00",
      "actor": "Admin"
    }
  ]
}
\end{lstlisting}

\subsubsection{Detailed Field Description Tables}

\textbf{Level 1 --- Root Document (Match/Project):}
\begin{table}[H]
\centering
\begin{tabular}{|l|l|l|p{5.5cm}|}
\hline
\textbf{Field} & \textbf{Type} & \textbf{Required} & \textbf{Description} \\ \hline
\_id & String & Yes & Custom ID format: \texttt{match\_<uuid8>} \\ \hline
name & String & Yes & Project/match name \\ \hline
date & String & Yes & Start date (ISO format) \\ \hline
date\_end & String & No & End date \\ \hline
status & String & Yes & Planned / In Progress / Done \\ \hline
created\_at & String & Yes & Creation timestamp (ISO) \\ \hline
categories & Array & Yes & Array of Phases (Level 2) \\ \hline
activity\_logs & Array & Yes & Array of activity log entries \\ \hline
\end{tabular}
\caption{Root Document Fields}
\end{table}

\textbf{Level 2 --- Category (Phase):}
\begin{table}[H]
\centering
\begin{tabular}{|l|l|p{7cm}|}
\hline
\textbf{Field} & \textbf{Type} & \textbf{Description} \\ \hline
name & String & Phase name (e.g., ``Logistics \& Venue'') \\ \hline
tasks & Array & Array of Tasks belonging to this phase \\ \hline
\end{tabular}
\caption{Category Fields}
\end{table}

\textbf{Level 3 --- Task:}
\begin{table}[H]
\centering
\begin{tabular}{|l|l|p{6cm}|}
\hline
\textbf{Field} & \textbf{Type} & \textbf{Description} \\ \hline
name & String & Task name \\ \hline
status & String & Todo / In Progress / Done / Incomplete \\ \hline
task\_type & String & logistics / personal / general \\ \hline
assigned\_to & String & Assignee name \\ \hline
assignee\_id & String & Reference ID to members collection \\ \hline
location & String & Execution location \\ \hline
cost & Float & Cost in USD \\ \hline
subtasks & Array & Array of Sub-tasks (Level 4) \\ \hline
\end{tabular}
\caption{Task Fields}
\end{table}

\textbf{Level 4 --- SubTask:}
\begin{table}[H]
\centering
\begin{tabular}{|l|l|p{6.5cm}|}
\hline
\textbf{Field} & \textbf{Type} & \textbf{Description} \\ \hline
name & String & Sub-task name \\ \hline
cost & Any & Cost --- can be numeric or string (requires Data Cleansing) \\ \hline
status & String & Status --- can be null or ``N/A'' (requires Data Cleansing) \\ \hline
\end{tabular}
\caption{SubTask Fields}
\end{table}

\subsubsection{Schema for the \texttt{members} Collection}
\begin{lstlisting}[language=json, caption=Document structure in the members collection]
{
  "_id": "mem_abc12345",
  "name": "Duong",
  "username": "duong",
  "password": "a665a459...7ae3",
  "role": "Member"
}
\end{lstlisting}

Passwords are hashed using SHA-256 before storage:
\begin{lstlisting}[language=Python, caption=Password hashing function]
import hashlib
def hash_password(password: str) -> str:
    return hashlib.sha256(password.encode()).hexdigest()
\end{lstlisting}

\subsection{Application-Level Validation Schema (Pydantic)}
The system uses Pydantic BaseModel to validate input data before writing to MongoDB, ensuring data integrity:

\begin{lstlisting}[language=Python, caption=models/schemas.py --- Pydantic Schema Definitions]
from pydantic import BaseModel, Field
from typing import Optional, List, Any

class SubTask(BaseModel):
    name: str
    cost: Optional[Any] = None    # Allow Any to test Data Cleansing
    status: Optional[str] = None  # Allow None to test Data Cleansing

class Task(BaseModel):
    name: str
    assigned_to: Optional[str] = None
    assignee_id: Optional[str] = None
    status: Optional[str] = "Todo"
    task_type: Optional[str] = "general"  # "logistics"|"personal"|"general"
    location: Optional[str] = None
    cost: Optional[float] = 0
    subtasks: Optional[List[SubTask]] = []

class Category(BaseModel):
    name: str
    tasks: Optional[List[Task]] = []

class MatchCreate(BaseModel):
    name: str
    date: str
    date_end: Optional[str] = None
    status: Optional[str] = "Planned"
    categories: Optional[List[Category]] = []
\end{lstlisting}

Notable design choice: the \texttt{cost} field of \texttt{SubTask} is declared as type \texttt{Any} instead of \texttt{float}. This is \textbf{intentional} to simulate real-world scenarios where users enter invalid values (e.g., the string ``N/A''). The Aggregation Pipeline discussed later handles these cases.


\subsection{Seed Data --- Populating Sample Data}

The file \texttt{scripts/seed\_data.py} automatically populates the database with 5 members and 10 matches. Each match contains 4 categories, 9 tasks, and 18 sub-tasks. In total, the system seeds \textbf{180 sub-tasks} randomly assigned across 5 members.

\begin{lstlisting}[language=Python, caption=Seeding sample data into MongoDB]
def seed():
    matches_collection = get_matches_collection()
    members_collection = get_members_collection()

    # Clear old data (overwrite)
    matches_collection.delete_many({})
    members_collection.delete_many({})

    # 1. Create 5 users
    users = [
        {"name": "Admin", "username": "admin", "role": "Admin"},
        {"name": "Duong", "username": "duong", "role": "Member"},
        {"name": "Vu", "username": "vu", "role": "Member"},
        {"name": "Tai", "username": "tai", "role": "Member"},
        {"name": "Dang", "username": "dang", "role": "Member"},
    ]
    for u in users:
        u["_id"] = f"mem_{uuid.uuid4().hex[:8]}"
        u["password"] = password_hash
    members_collection.insert_many(users)

    # 2. Create 10 matches with categories + activity_logs
    for i in range(1, 11):
        match = {
            "_id": f"match_{uuid.uuid4().hex[:8]}",
            "name": f"Friendly match round {i}",
            "date": (datetime.now() + timedelta(days=i*2)).strftime("%Y-%m-%d"),
            "status": "Planned",
            "categories": categories,    # 4 categories, 9 tasks, 18 subtasks
            "activity_logs": [{
                "action": f"Created new match round {i}",
                "timestamp": datetime.now().isoformat(),
                "actor": "Admin"
            }],
            "created_at": datetime.now().isoformat()
        }
    matches_collection.insert_many(matches)
\end{lstlisting}

Key point: each task is randomly assigned to a member using \texttt{random.choice(user\_names)}, creating a realistic workload distribution for testing the Aggregation Pipeline later.

\subsection{CRUD Operations on Nested Documents}

\subsubsection{CREATE --- Creating a New Project}
When creating a new match, if no categories are provided, the system automatically loads a default template \texttt{DEFAULT\_CATEGORIES} containing 5 phases: Logistics \& Venue, Finance \& Team Fund, Tactics \& Strategy, Media, and Personal Tasks.

\begin{lstlisting}[language=Python, caption=Create new project --- POST /match]
@router.post("/")
def create_match(payload: MatchCreate):
    import uuid
    new_id = f"match_{uuid.uuid4().hex[:8]}"
    doc = payload.dict()
    if not doc.get("categories"):
        doc["categories"] = DEFAULT_CATEGORIES  # 5 default phases
    doc["_id"] = new_id
    doc["created_at"] = datetime.now().isoformat()
    doc["activity_logs"] = [{
        "action": f"Created new match: {payload.name}",
        "timestamp": datetime.now().isoformat(),
        "actor": "Admin"
    }]
    matches_collection.insert_one(doc)
    return {"message": "Match created", "id": new_id}
\end{lstlisting}

\subsubsection{READ --- Single Query Retrieval}
This is the core feature: \textbf{only 1 single query} is needed to fetch the entire 4-level data tree. The response time (in milliseconds) is measured and returned to the frontend.

\begin{lstlisting}[language=Python, caption=Read entire project --- GET /match/\{match\_id\}]
@router.get("/{match_id}")
def get_match(match_id: str):
    start_time = time.time()
    # === ONLY 1 SINGLE QUERY ===
    match = matches_collection.find_one({"_id": match_id})
    if not match:
        raise HTTPException(status_code=404, detail="Match not found")
    match["id"] = str(match["_id"])
    del match["_id"]
    response_time_ms = round((time.time() - start_time) * 1000, 2)
    return {
        "data": match,
        "metadata": {
            "response_time_ms": response_time_ms,
            "query_count": 1,
            "message": "Fetched entire nested document with 1 single MongoDB query"
        }
    }
\end{lstlisting}

Actual results: response time is typically only \textbf{2--5ms}, while a SQL approach with 4 JOIN statements can take 50--200ms for the same data volume.

The project listing API (\texttt{GET /match}) uses \textbf{projection} to exclude heavy fields:
\begin{lstlisting}[language=Python, caption=Projection to exclude heavy fields when listing]
matches = list(matches_collection.find({}, {"categories": 0, "activity_logs": 0}))
\end{lstlisting}

\subsubsection{UPDATE --- Updating Deeply Nested Tasks with Array Filters}
The biggest challenge of Nested Documents is updating an element buried 3 levels deep. MongoDB provides \texttt{array\_filters} to solve this:

\begin{lstlisting}[language=Python, caption=Update deeply nested task --- POST /match/\{id\}/update-task]
@router.post("/{match_id}/update-task")
def update_task(match_id: str, payload: UpdateTaskRequest):
    set_ops = {
        "categories.$[c].tasks.$[t].status": payload.new_status
    }
    if payload.location is not None:
        set_ops["categories.$[c].tasks.$[t].location"] = payload.location
    if payload.assigned_to is not None:
        set_ops["categories.$[c].tasks.$[t].assigned_to"] = payload.assigned_to
    if payload.cost is not None:
        set_ops["categories.$[c].tasks.$[t].cost"] = payload.cost
    if payload.subtasks is not None:
        set_ops["categories.$[c].tasks.$[t].subtasks"] = [
            s.dict() for s in payload.subtasks
        ]
    result = matches_collection.update_one(
        {"_id": match_id},
        {"$set": set_ops, "$push": {"activity_logs": log_entry}},
        array_filters=[
            {"c.name": payload.category_name},
            {"t.name": payload.task_name}
        ]
    )
\end{lstlisting}

The syntax \texttt{[c]} and \texttt{[t]} are \textbf{positional filtered operators} --- MongoDB traverses the \texttt{categories} array to find the element whose \texttt{name} matches \texttt{c.name}, then within the child \texttt{tasks} array finds the element matching \texttt{t.name}. This operation executes at the database engine level, achieving $OO(1) performance without loading the document into application memory.

\subsubsection{UPDATE --- Saving Tree Structure After Drag-and-Drop}
When users drag-and-drop to reorder phases/tasks on the Tree interface, the entire new \texttt{categories} array is sent and overwritten:

\begin{lstlisting}[language=Python, caption=Save tree structure --- POST /match/\{id\}/update-tree]
@router.post("/{match_id}/update-tree")
def update_tree(match_id: str, payload: UpdateTreeRequest):
    result = matches_collection.update_one(
        {"_id": match_id},
        {
            "$set": {"categories": payload.categories},
            "$push": {"activity_logs": {
                "action": "Rearranged tree structure (drag-and-drop)",
                "timestamp": datetime.now().isoformat(),
                "actor": payload.actor or "Admin"
            }}
        }
    )
\end{lstlisting}

\subsubsection{DELETE --- Deleting a Project}
Only allows deletion of matches that are not yet completed (\texttt{status != "Done"}):
\begin{lstlisting}[language=Python, caption=Delete project --- DELETE /match/\{id\}]
@router.delete("/{match_id}")
def delete_match(match_id: str):
    match = matches_collection.find_one({"_id": match_id})
    if match.get("status") == "Done":
        raise HTTPException(status_code=400,
            detail="Cannot delete a completed match")
    matches_collection.delete_one({"_id": match_id})
\end{lstlisting}

\subsection{Activity Log Mechanism --- Atomic Logging with \texttt{push}}

The \texttt{activity\_logs} array is stored directly in the root document. Every status change operation uses the \texttt{push} operator combined with \texttt{set} in \textbf{the same \texttt{update\_one} call}, ensuring atomicity:

\begin{lstlisting}[language=Python, caption=Structure of a log entry]
log_entry = {
    "action": "[Logistics] 'Book the pitch' -> 'Done' | 50 USD | Duong",
    "timestamp": "2026-05-02T15:30:00",
    "actor": "Duong"
}
\end{lstlisting}

The API \texttt{GET /match/\{id\}/logs} retrieves logs and returns them in newest-first order:
\begin{lstlisting}[language=Python, caption=Activity Log retrieval API]
@router.get("/{match_id}/logs")
def get_logs(match_id: str):
    match = matches_collection.find_one(
        {"_id": match_id}, {"activity_logs": 1}  # Projection: only logs
    )
    logs = match.get("activity_logs", [])
    return {"logs": list(reversed(logs)), "total": len(logs)}
\end{lstlisting}

The system also features \textbf{automatic status transitions} based on real-time: the function \texttt{\_auto\_update\_match\_status()} compares the current time with \texttt{date}/\texttt{date\_end} to automatically transition Planned {\textrightarrow} In Progress {\textrightarrow} Done, while logging with actor set to ``System''.

\subsection{Automatic Incomplete Task Marking}
When a match transitions to In Progress or Done, the function \texttt{\_move\_unfinished\_tasks\_to\_incomplete()} traverses the entire categories tree and marks all tasks/subtasks with status Todo or In Progress as ``Incomplete'':

\begin{lstlisting}[language=Python, caption=Automatically marking unfinished tasks]
def _move_unfinished_tasks_to_incomplete(match_id, actor):
    match = matches_collection.find_one({"_id": match_id})
    categories = match.get("categories", [])
    for cat in categories:
        for task in cat.get("tasks", []):
            if task.get("status") in ("Todo", "In Progress"):
                task["status"] = "Incomplete"
                for st in task.get("subtasks", []):
                    if st.get("status") in ("Todo", "In Progress"):
                        st["status"] = "Incomplete"
    matches_collection.update_one(
        {"_id": match_id},
        {"$set": {"categories": categories},
         "$push": {"activity_logs": log_entry}}
    )
\end{lstlisting}


\subsection{Aggregation Pipeline --- Performance Scoring \& Data Cleansing}

The API \texttt{GET /match/\{id\}/analytics} runs \textbf{4 Aggregation Pipelines} to perform comprehensive data analysis. All computation logic executes at the database engine level, without loading data into application RAM.

\subsubsection{Pipeline 1: Member Performance Scoring}

\textbf{Step 1 --- \texttt{unwind}:} Flatten nested arrays into a document stream:
\begin{lstlisting}[language=Python, caption=Unwind categories and tasks]
{"$match": {"_id": match_id}},
{"$unwind": "$categories"},
{"$unwind": "$categories.tasks"},
\end{lstlisting}

\textbf{Step 2 --- Data Cleansing with \texttt{filter}:} Automatically detect and exclude sub-tasks with invalid data. Filter conditions:
\begin{itemize}
    \item \texttt{status} must be in the valid list: ``Todo'', ``In Progress'', ``Done'', ``Incomplete''. Values like \texttt{null}, \texttt{""}, \texttt{"N/A"} are excluded.
    \item \texttt{cost} must be a numeric type (\texttt{int}, \texttt{double}, \texttt{long}) or \texttt{null}. String values (e.g., \texttt{"N/A"}, \texttt{"pending"}) are excluded to prevent crashes during sum calculations.
\end{itemize}

\begin{lstlisting}[language=Python, caption=Data Cleansing --- filtering valid subtasks]
{"$addFields": {
    "valid_subtasks": {
        "$filter": {
            "input": {"$ifNull": ["$categories.tasks.subtasks", []]},
            "as": "st",
            "cond": {"$and": [
                {"$in": ["$$st.status",
                    ["Todo", "In Progress", "Done", "Incomplete"]]},
                {"$or": [
                    {"$eq": [{"$type": "$$st.cost"}, "int"]},
                    {"$eq": [{"$type": "$$st.cost"}, "double"]},
                    {"$eq": [{"$type": "$$st.cost"}, "long"]},
                    {"$eq": [{"$type": "$$st.cost"}, "null"]},
                    {"$not": {"$gt": ["$$st.cost", None]}}
                ]}
            ]}
        }
    },
    "all_subtasks": {"$ifNull": ["$categories.tasks.subtasks", []]}
}}
\end{lstlisting}

The system counts skipped sub-tasks (\texttt{skipped\_subtask\_count}) via \texttt{size(all) - size(valid)}, returning this count to the frontend for transparency.

\textbf{Step 3 --- Scoring with \texttt{switch}:}
\begin{lstlisting}[language=Python, caption=Performance scoring rules]
"task_score": {"$switch": {"branches": [
    {"case": {"$eq": ["$categories.tasks.status", "Done"]}, "then": 10},
    {"case": {"$eq": ["$categories.tasks.status", "In Progress"]}, "then": 5},
    {"case": {"$eq": ["$categories.tasks.status", "Todo"]}, "then": 1}
], "default": 0}},
"subtask_bonus": {"$add": [
    {"$multiply": ["$subtask_done_count", 3]},
    {"$multiply": ["$subtask_inprogress_count", 1]}
]}
\end{lstlisting}

Scoring rules table:
\begin{table}[H]
\centering
\begin{tabular}{|l|c|l|}
\hline
\textbf{Type} & \textbf{Points} & \textbf{Description} \\ \hline
Task Done & 10 & Completed main task \\ \hline
Task In Progress & 5 & Currently working \\ \hline
Task Todo & 1 & Task accepted \\ \hline
Sub-task Done (bonus) & 3 & Completed sub-task \\ \hline
Sub-task In Progress (bonus) & 1 & Working on sub-task \\ \hline
\end{tabular}
\caption{Performance scoring rules}
\end{table}

\textbf{Step 4 --- \texttt{group} by member:}
\begin{lstlisting}[language=Python, caption=Group by assigned\_to]
{"$group": {
    "_id": "$categories.tasks.assigned_to",
    "task_score": {"$sum": "$task_score"},
    "subtask_bonus": {"$sum": "$subtask_bonus"},
    "total_score": {"$sum": {"$add": ["$task_score", "$subtask_bonus"]}},
    "total_tasks": {"$sum": 1},
    "done_tasks": {"$sum": "$is_done"},
    "total_subtasks": {"$sum": "$valid_subtask_count"},
    "done_subtasks": {"$sum": "$subtask_done_count"},
    "skipped_subtasks": {"$sum": "$skipped_subtask_count"}
}}
\end{lstlisting}

\textbf{Step 5 --- Calculating completion rate:}
\begin{lstlisting}[language=Python, caption=Completion rate with division-by-zero prevention]
{"$addFields": {
    "completion_rate": {"$round": [
        {"$multiply": [
            {"$divide": ["$done_tasks", {"$max": ["$total_tasks", 1]}]},
            100
        ]}, 1
    ]}
}}
\end{lstlisting}
Uses \texttt{max[total\_tasks, 1]} to prevent division-by-zero errors when a member has no assigned tasks.

\subsubsection{Pipeline 2: Total Fund Calculation}
Only sums the cost of sub-tasks where \texttt{cost} is a numeric type:
\begin{lstlisting}[language=Python, caption=Fund calculation pipeline]
fund_pipeline = [
    {"$match": {"_id": match_id}},
    {"$unwind": "$categories"},
    {"$unwind": "$categories.tasks"},
    {"$unwind": {"path": "$categories.tasks.subtasks",
                 "preserveNullAndEmptyArrays": True}},
    {"$match": {
        "categories.tasks.subtasks.cost": {"$type": ["int", "double", "long"]}
    }},
    {"$group": {"_id": None,
                "total_fund": {"$sum": "$categories.tasks.subtasks.cost"}}}
]
\end{lstlisting}

\subsubsection{Pipeline 3: Overall Progress}
\begin{lstlisting}[language=Python, caption=Progress calculation pipeline]
progress_pipeline = [
    {"$match": {"_id": match_id}},
    {"$unwind": "$categories"},
    {"$unwind": "$categories.tasks"},
    {"$match": {"categories.tasks.status": {
        "$exists": True,
        "$nin": [None, "null", "", "N/A", "Todo"]
    }}},
    {"$group": {"_id": None,
        "total": {"$sum": 1},
        "done": {"$sum": {"$cond": [
            {"$eq": ["$categories.tasks.status", "Done"]}, 1, 0]}}
    }}
]
\end{lstlisting}

\subsubsection{Pipeline 4: Counting Skipped Sub-tasks}
\begin{lstlisting}[language=Python, caption=Pipeline to count invalid sub-tasks]
skipped_pipeline = [
    {"$match": {"_id": match_id}},
    {"$unwind": "$categories"},
    {"$unwind": "$categories.tasks"},
    {"$unwind": {"path": "$categories.tasks.subtasks",
                 "preserveNullAndEmptyArrays": False}},
    {"$match": {"$or": [
        {"categories.tasks.subtasks.status": {"$in": [None, "null", "", "N/A"]}},
        {"categories.tasks.subtasks.status": {"$exists": False}},
        {"categories.tasks.subtasks.cost": {"$type": "string"}}
    ]}},
    {"$count": "total_skipped"}
]
\end{lstlisting}

The API returns comprehensive results including the leaderboard, total fund, progress, and the number of skipped invalid sub-tasks --- fully transparent to the user.

\newpage

%% === SECTION 3: CONCLUSION ===
\section{Conclusion}
This report has presented a detailed design and implementation of the MongoDB database for the Creative Project Portfolio platform. Key achievements:

\begin{enumerate}
    \item \textbf{4-Level Nested Document Schema:} Encapsulates the entire Project {\textrightarrow} Phase {\textrightarrow} Task {\textrightarrow} Sub-task structure in a single document, maximizing MongoDB's Data Locality.
    \item \textbf{Single Query Performance:} Retrieves the complete data tree with just 1 \texttt{find\_one()} call, achieving 2--5ms response times --- significantly outperforming multi-JOIN SQL approaches.
    \item \textbf{Atomic Updates with Array Filters:} Precisely updates elements nested 3 levels deep without loading the document into application memory, simultaneously recording Activity Logs via \texttt{push}.
    \item \textbf{Advanced Aggregation Pipelines:} 4 pipelines running in parallel for member performance analysis, fund calculation, progress measurement, and invalid data counting --- all at the database engine level.
    \item \textbf{Automatic Data Cleansing:} Uses \texttt{filter} + \texttt{type} to detect and skip sub-tasks with null/``N/A'' status or string-type cost values, preventing crashes in computation pipelines.
\end{enumerate}

%% === SECTION 4: DISCUSSION ===
\section{Discussion and Future Work}
\subsection{Advantages}
\begin{itemize}
    \item Extremely high read performance thanks to Data Locality.
    \item Flexible schema, easy to extend fields without migrations.
    \item Aggregation Pipeline processes entirely at the DB level, saving application RAM.
\end{itemize}

\subsection{Limitations}
\begin{itemize}
    \item MongoDB's 16MB document size limit. Extremely large projects (hundreds of thousands of sub-tasks) may hit this limit.
    \item Updating deeply nested elements is more complex compared to SQL \texttt{UPDATE ... WHERE}.
\end{itemize}

\subsection{Future Directions}
\begin{itemize}
    \item \textbf{Data Sharding:} Separate attached resources (files, images) into a dedicated collection when documents grow too large.
    \item \textbf{Redis Cache:} Cache JSON responses to achieve sub-millisecond response times.
    \item \textbf{Real-time Sync:} Implement MongoDB Change Streams + WebSocket to synchronize the Kanban board in real-time across all connected users.
    \item \textbf{Indexing:} Create indexes on \texttt{categories.tasks.assigned\_to} and \texttt{status} to accelerate Aggregation queries.
\end{itemize}



\end{document}
